Accessible, realistic genome simulation with selection using \stdpopsim

### 2025-03-16 analysis2_manuscript.tex
plr edits to rest of paper

    % stdpopsim
    In light of the complexities associated with estimating and simulating
    population genetic models, it is no surprise that
    a lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers \citep[e.g.,][]{ragsdale2020lessons}.
    This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community 
    \citep[e.g.,][]{speidel2021inferring, wang2021automatic, yang2022trustgwas, degiorgio2022spatially, browning2023fast, schweiger2023ultrafast, temple2024modeling, haag2025pandora}.
    Yet, previous releases of \stdpopsim did not include
    models of selection, which is a major limitation for more empirically motivated
    applications of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding processes such
    as adaptation \citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps 
    \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, 
    and the impact of background selection on genetic diversity
    \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.
    Ideally, one would like
    to have a simple, unified framework for simulating both neutral and
    non-neutral evolutionary processes and to be able to compare the
    results of these simulations to empirical data in a manner that is
    both accessible to a wide range of researchers and highly reproducible. 
    Furthermore, the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the size and density of functional elements, and
    non-equilibrium demographic histories. Moreover, one should be able
    to model all of these features simultaneously using estimates and annotations obtained
    from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.

@@

    As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on specices' publicly available functional genomic elements 
    and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
    By using annotation objects, one can, for example, simulate populations where selected mutations (whose selection
    coefficients are drawn from a specified DFE) occur within annotation regions (e.g. exons), while all mutations
    occurring outside of these regions do not affect fitness.
    There are relatively few species with published DFE estimates in the literature.
    \stdpopsim currently has implemented DFEs for five species in the catalogue:
    %% MISSING SOME CITATIONS IF WE"RE AIMING TO HAVE ALL OF THEM IN HERE, I THINK?
    \textit{Arabidopsis thaleana} \citep{huber2018gene},
    \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
    humans \citep{huber2017determining,kim2017inference},
    \textit{Mus musculus} \citep{booker2021selective},
    and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While the DFEs are from a limited number of species, these DFEs can be applied to other species
    in the catalog if and when it is suitable. 
    For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Indeed, the DFE labeled \texttt{HomSap/Mixed_K23} is recommended by \citet{kyriazis2022using}
    for ``general'' use across species (lacking more specific information).
    Furthermore, the user can specify a custom DFE and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have 
    a published DFE included in the catalog.

@@

    We then applied each method to data simulated with sweeps for two samples (YRI and CEU)
    in each of 100 locations along human chromosome 1 (Figure \ref{fig:chr1_power}).
    First and most strikingly, the power to detect sweeps varies considerably along the chromosome,
    and this variation is much greater than the variation in power between CEU and YRI or between
    null models. This is likely due to the effects of recombination rate heterogeniety (Figure \ref{fig:chr1_power}).
    Next, we notice that power to detect sweeps is slightly lower when the test statistic threshold
    is determined using simulated data with background selection (BGS null model).
    %This observation is consistent with what was previously shown by \cite{schrider2020background},
    %but the effect is relatively minor. DRS: cutting this because my paper showed that power decreases when our test set has BGS and our training set does not, so not quite the same.
    This is likely because the strength of selection acting on exons is relatively weak, and the
    sweep detection methods we have used are relatively robust to the effects of background selection. 
    Finally, the power to detect sweeps is generally higher in the YRI sample
    than in the CEU sample, which is consistent with the fact that YRI has a larger ancestral effective population size and
    thus both stronger selection and more variation to detect a sweep \citep[e.g.,][]{simonsen1995properties}.
    Conversely, the CEU population experienced the out-of-Africa bottleneck that may mute the signal of a sweep.

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=0.8 \textwidth]{figures/sweeps/relationship_power_cM.pdf}
        \caption{
        Power to detect selective sweeps 
        %IG: shortened the title line
        %at 100 locations along human
        %chromosome 1 under a three population out-of-Africa model
        %\citep{gutenkunst2009inferring}  when using null models with and
        %without purifying selection
        as a function of local recombination rate.
        This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
        but with the 100 genomic segments sorted according to their
        average recombination rates instead of position along chromosome 1.
    	Genomic segments were simulated with sweeps under a three population out-of-Africa model
        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
        Three methods for detecting sweeps were applied to simulated data:
        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
        2) \diploshic \citep{kern2018diplos} (middle row),
        and 3) reduced diversity ($\pi$) (bottom row).
        Power (true positive rate) is shown for these methods for the CEU and YRI
        samples (left and right respectively).
        The thresholds of the test statistics were set to control for
        $5\%$ false positive rate under a neutral null model (blue)
        and a null model with background selection from deleterious mutations in exons (red).
        Fitted lines represent loess smoothed regressions.
        }
        \label{fig:power-recomb}
    \end{figure}

@@

    \subsection*{Whole genome simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using the \stdpopsim{} framework. We ran replicate datasets for each genome/parameter combination,
    in which chromosomes were simulated separately from one another.
    All simulations were executed using the \slim engine with a burn-in of $10N$ generations
    (where $N$ is the initial population size), and a scaling factor of 1 (no scaling).
    All simulations were implemented using as part of a \snakemake workflow,
    where the code utilized is similar to the one shown in Figure 1B.
    The complete simulation workflow is available at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/simulation.snake}
    (see \textbf{Data availablity} below).

    \subsubsection*{Simulations of complete human genomes}
    Simulation of human genomes used a genetic map from the HapMap Project
    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
    the out of Africa demographic model with archaic admixture from
    \citet{ragsdale2019models} (\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}).
    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
    \citet{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
    (\citet{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
    Each simulation generated 100 diploid genomes from each of the three extant human populations (YRI, CEU, and CHB),
    with each genome consisting of 22 autosomes.
    Using these settings, we generated three replicate datasets without selection
    and three replicate datasets with selection.
    Three additional datasets, used in Figure \ref{fig:homsap-dfe.constant},
    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $10,000$.
    Each of these simulations generated 100 diploid genomes.
    The configuration file used in combination with our \snakemake workflow
    for these simulations can be found at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_HomSap.yml}.

    \subsubsection*{Simulations of complete vaquita porpoise genomes}
    Simulation of vaquita porpoise genomes used 
    the two-epoch demographic model from \citet{robinson2022critically} (\stdpopsim label \texttt{Vaquita2Epoch\_1R22}).
    Since no genetic map is currently available for this species, we assumed a constant recombination rate of  $10^{-8}$
    across all chromosomes \citep{morin2021}.
    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
    \citet{robinson2022critically} (\stdpopsim label \texttt{Gamma\_R22}) to exons annotated from the vaquita genome assembly
    (\citet{morin2021}; \stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}).
    This DFE model implements a relationship between selection and dominance as follows:
    very deleterious mutations (with $s<-0.1$) were set to be recessive ($h=0$),
    mutations with $s\in [-0.1,-0.01)$ were set to have a dominance coefficient of $h=0.01$,
    mutations with $s\in [-0.01,-0.001)$ were set to have a dominance coefficient of $h=0.1$,
    and very mild mutations (with $s\in [-0.001,0)$) were set to have a nearly additive effect ($h=0.4$).
    This approach is similar to what was used in the simulations conducted by \textcite{robinson2022critically}
    and is easily implemented in \stdpopsim's \texttt{DFE} class.
    Each simulation generate 100 diploid genomes,
    with each genome consisting of 21 autosomes.
    Using these settings, we generated three replicate datasets without selection
    and three replicate datasets with selection.
    Three additional datasets, used in Figure \ref{fig:vaquita-dfe.constant},
    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $3,500$.
    \igcomment{If we end up using Fig. S3, then we should add a line describing it's simulation here.}
    The configuration file used in combination with our \snakemake workflow
    for these simulations can be found at 
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_PhoSin.yml}.

@@

    % ADK: I've added a section below that describes the DFE inference methods used in this study.
    For comparison among methods, estimates were rescaled. 
    As described above, each DFE-inference method scales selection coefficients slightly differently.
    We thus rescaled each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
    The output of \grapes  was scaled by a factor of  $\mu / \theta$,
    and the output of \polydfe was saled by $2 \mu / \theta$, 
    where $\mu$ is the assumed mutation rate and 
    $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
    For \dadicli, the gamma distribution is parameterized using the shape and rate ($\alpha$ and $\nu$, respectively). 
    The mean value of the distribution, which is $s$ is transformed from $\alpha$ and $\nu$ and rescaled as $\alpha * \nu / (\theta / (4 \mu L p))$,
    where $L$ is the total number of sites available for the analysis,
    and $p$ is the proportion of sites under selection.
    Complete details for the DFE inference analysis can be found in the
    \snakemake workflow at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/dfe.snake}.

@@

    Each of the three sweep-detection methods is based on a test statistic.
    Reduced nucleotide diversity is based on nucleotide diversity  ($\pi$) 
    computed directly from the underlying tree sequences using \tskit \citep{ralph2020efficiently}
    over 10 equally sized and non-overlapping subwindows.
    \sweepfinder \citep{degiorgio2016sweepfinder2} is based on a composite likelihood ratio (CLR) 
    computed at 21 uniformly distributed positions along the simulated region.
    \diploshic \citep{kern2018diplos} uses supervised learning to train a sweep classifier.
    Its training phase used \texttt{discoal} \citep{kern2016discoal} to simulate 20 haploid sequences of length 1.21Mbp
    with a constant population size of $N_e=10,000$, a mutation rate of $2\times 10^{-8}$, and a recombination rate of $2\times 10^{-8}$.
    For \texttt{discoal} simulations that included selective sweeps,
    selection coefficients were drawn from a uniform distribution between $0.001$ and $0.10$,
    and the time of the sweep was chosen from a uniform distribution between $0$ and $0.01N_e$ generations in the past.
    For soft sweeps, the frequency at which selection was introduced was drawn from a uniform distribution between $0.0001$ and $0.1$.
    We note that since the training set of \diploshic was generated under a simple model with
    constant mutation and recombination rates and constant effective population size,
    we expect it to perform suboptimally on the simulated segments of human chromosome 1,
    which were generated using more relaistic simulations (see \textbf{Simulations with selective sweeps} above).

    The three sweep-detection methods were associated with thresholds for their respective test statistics
    ($\pi$, \sweepfinder's CLR, or \diploshic's probability),
    which were determined using an empirical distribution representing a null model without sweeps.
    We considered two null models: a neutral evolution model (neutral),
    and a model with background selection from deleterious mutations in exons (BGS).
    Each null model was represented by a different set of $100,000$ simulations generated according to the appropriate model
    (see \textbf{Simulations with selective sweeps} above).
    Thus, for each test statistic we determined two thresholds corresponding to these two different null models (neutral and BGS)
    using the extreme $5\%$ quantile in each empirical distribution.
    Simulated regions where the central 1Mb had a recombination rate of less than $10^{-3}cM$ were excluded from the analysis.
    This excluded 7 regions from the analysis. 
    Then, power (or true positive rate) to detect a sweep by a given method was determined for
    100 evenly-spaced positions on chromosome 1 as follows:
    each sweep-detection method was applied to the 1,000 
    simulated datasets generated with a hard sweep (neutral and BGS) for that position,
    and the power was set to be the proportion of simulations (out of 1,000)
    in which the test statistic exceeded the threshold determined by each of the two null models (neutral and BGS).
    We note that using an empirical distribution to determine the threshold for calling a selective
    sweep with \diploshic deviates from the original implementation of the method,
    however for consistency among methods we chose to take this approach.
    The complete \snakemake workflow for sweep detection can be found at
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/sweep_simulate.snake}.

@@

\section*{Funding}
    \label{funding}
    MFR, ST, NP, and ADK were supported in part by NIH award R35GM148253.
    RNG and LNT were supported in part by NIH award R35GM149235.
    DRS was supported in part by NIH award R35GM138286.
    KEL was supported by NIH award R35GM119856.

    %other funding sources?
\printbibliography

@@

%sweep power vs exon density
\begin{figure}
    \centering
    \includegraphics[width=0.8 \textwidth]{figures/sweeps/relationship_power_exon.pdf}
    \caption{
    %IG: copied new caption from fig 6 and adjusted.
        Power to detect selective sweeps as a function of local exon density.
        This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
        but with the 100 genomic segments sorted according to their
        exon density (percentage of exonic bases per window)instead of position along chromosome 1.
	    Genomic segments were simulated with sweeps under a three population out-of-Africa model
        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
        Three methods for detecting sweeps were applied to simulated data:
        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
        2) \diploshic \citep{kern2018diplos} (middle row),
        and 3) reduced diversity ($\pi$) (bottom row).
        Power (true positive rate) is shown for these methods for the CEU and YRI
        samples (left and right respectively).
        The thresholds of the test statistics were set to control for
        $5\%$ false positive rate under a neutral null model (blue)
        and a null model with background selection from deleterious mutations in exons (red).
        Fitted lines represent loess smoothed regressions.
    }
    \label{fig:power-exon}
\end{figure}

>>>>

    % stdpopsim
    In light of the complexities associated with estimating and simulating
    population genetic models, it is no surprise that
    a lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers \citep[e.g.,][]{ragsdale2020lessons}.
    This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community 
    \citep[e.g.,][]{speidel2021inferring, wang2021automatic, yang2022trustgwas, degiorgio2022spatially, browning2023fast, schweiger2023ultrafast, temple2024modeling, haag2025pandora}.
    Yet, previous releases of \stdpopsim did not include
    models of selection, which is a major limitation for more empirically motivated
    applications of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding processes such
    as adaptation \citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps 
    \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, 
    and the impact of background selection on genetic diversity
    \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.
    Ideally, one would like
    to have a simple, unified framework for simulating both neutral and
    non-neutral evolutionary processes and to be able to compare the
    results of these simulations to empirical data in a manner that is
    both accessible to a wide range of researchers and highly reproducible. 
    Furthermore, the framework should include the complex realities of 
    genomes, including heterogeneous recombination rates, 
    variation in the size and density of functional elements, and
    non-equilibrium demographic histories. Moreover, one should be able
    to model all of these features simultaneously using estimates and annotations obtained
    from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.

@@

    As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on species' publicly available functional genomic elements 
    and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
    By using annotation objects, one can, for example, simulate populations where selected mutations (whose selection
    coefficients are drawn from a specified DFE) occur within annotation regions (e.g. exons), while all mutations
    occurring outside of these regions do not affect fitness.
    There are relatively few species with published DFE estimates in the literature.
    \stdpopsim currently has implemented DFEs for five species in the catalogue:
    \textit{Arabidopsis thaliana} \citep{huber2018gene},
    \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining,zhen2021greater},
    humans \citep{huber2017determining,kim2017inference,kyriazis2022using,castellano2019compmarison,castellano2019compmarison,rodrigues2024shared},
    \textit{Mus musculus} \citep{booker2021selective},
    and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While the DFEs are from a limited number of species, these DFEs can be applied to other species
    in the catalog if and when it is suitable. 
    For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Indeed, the DFE labeled \texttt{HomSap/Mixed\_K23} is recommended by \citet{kyriazis2022using}
    for ``general'' use across species (lacking more specific information).
    Furthermore, the user can specify a custom DFE and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have 
    a published DFE included in the catalog.

@@

    We then applied each method to data simulated with sweeps for two samples (YRI and CEU)
    in each of 100 locations along human chromosome 1 (Figure \ref{fig:chr1_power}).
    First and most strikingly, the power to detect sweeps varies considerably along the chromosome,
    and this variation is much greater than the variation in power between CEU and YRI or between
    null models. This is likely due to the effects of recombination rate heterogeneity (Figure \ref{fig:chr1_power}).
    Next, we notice that power to detect sweeps is slightly lower when the test statistic threshold
    is determined using simulated data with background selection (BGS null model).
    %This observation is consistent with what was previously shown by \cite{schrider2020background},
    %but the effect is relatively minor. DRS: cutting this because my paper showed that power decreases when our test set has BGS and our training set does not, so not quite the same.
    This is likely because the strength of selection acting on exons is relatively weak, and the
    sweep detection methods we have used are relatively robust to the effects of background selection. 
    Finally, the power to detect sweeps is generally higher in the YRI sample
    than in the CEU sample, which is consistent with the fact that YRI has a larger ancestral effective population size and
    thus both stronger selection and more variation to detect a sweep \citep[e.g.,][]{simonsen1995properties}.
    Conversely, the CEU population experienced the out-of-Africa bottleneck that may mute the signal of a sweep.

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=0.8 \textwidth]{figures/sweeps/relationship_power_cM.pdf}
        \caption{
        Power to detect selective sweeps 
        %IG: shortened the title line
        %at 100 locations along human
        %chromosome 1 under a three population out-of-Africa model
        %\citep{gutenkunst2009inferring}  when using null models with and
        %without purifying selection
        as a function of local recombination rate.
        This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
        but with the 100 genomic segments plotted against their
        average recombination rates instead of position along chromosome 1.
    	Genomic segments were simulated with sweeps under a three population out-of-Africa model
        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
        Three methods for detecting sweeps were applied to simulated data:
        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
        2) \diploshic \citep{kern2018diplos} (middle row),
        and 3) reduced diversity ($\pi$) (bottom row).
        Power (true positive rate) is shown for these methods for the CEU and YRI
        samples (left and right respectively).
        The thresholds of the test statistics were set to obtain a
        $5\%$ false positive rate under a neutral null model (blue)
        and a null model with background selection from deleterious mutations in exons (red).
        Fitted lines represent loess smoothed regressions.
        }
        \label{fig:power-recomb}
    \end{figure}

@@

    \subsection*{Annotations}
    Currently available annotation data in \stdpopsim were downloaded from Ensembl.
    Coding sequence and exon annotations are available for
    \textit{Arabidopsis thaliana} \citep{cheng2017araport},
    \textit{Drosophila melanogaster} \citep{hoskins2015release},
    \textit{Homo sapiens} \citep{hunt2018ensembl},
    and \textit{Phocoena sinus} \citep[vaquita porpoise,][]{morin2021}.

    \subsection*{Quality control and software development}
    Following the protocol established previously for other aspects of \stdpopsim{} \citep{adrion2020community},
    new DFEs are subjet to a rigorous quality control (QC) process,
    by which a separate individual re-implements the same DFE, and the two implementations
    are programatically tested for equality.
    The QC process identified a substantial bug in one DFE:
    the \textit{Drosophila melanogaster} DFE \texttt{LognormalPlusPositive\_R16} \citet{booker2021selective}.
    Due to an error in coding the log-normal distribution, deleterious selection coefficients
    were several orders of magnitude too large.
    The bug affected results in a separate project from one of our groups
    that made use of this DFE before QC had been done.
    All changes and additions to the code must pass a comprehensive set of unit tests:
    at time of writing, the library includes over 11,000 lines of testing code,
    that run 2,862 unit tests, including comparisons to QC implementations
    and cross-checks for agreement, including between \slim and \msprime implementations of the same models.
    For instance, for testing we make use of \slim's scriptability to separately record in simulation metadata
    information about current population sizes, DFEs, etcetera,
    which are then programmatically compared to the models used to generate the \slim script
    that ran the simulation.
    We make use of automatic test ``code coverage'' tools to help ensure that all portions of the code base
    are validated by these unit tests.

    \subsection*{Whole genome simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using \stdpopsim{}. We ran replicate datasets for each genome/parameter combination,
    in which chromosomes were simulated separately from one another.
    All simulations were executed using the \slim engine with a burn-in of $10N$ generations
    (where $N$ is the initial population size), and a scaling factor of 1 (no scaling).
    All simulations were implemented using a \snakemake workflow,
    where the code utilized is similar to that shown in Figure~\ref{fig:schematic}B.
    The complete simulation workflow is available at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/simulation.snake}
    (see \textbf{Data availability} below).

    \subsubsection*{Simulations of complete human genomes}
    Simulation of human genomes used a genetic map from the HapMap Project
    \citep[\stdpopsim label \texttt{HapMapII\_GRCh38}]{international2007second}  and
    the out of Africa demographic model with archaic admixture from
    \citet[\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}]{ragsdale2019models}.
    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
    \citet[\stdpopsim label \texttt{Gamma\_K17}]{kim2017inference}  to exons annotated by the HAVANA group release 104
    \citep[\stdpopsim label \texttt{ensembl\_havana\_104\_exons}]{ensembl2018}.
    Each simulation generated 100 diploid genomes from each of three extant human populations (labeled YRI, CEU, and CHB),
    with each genome consisting of 22 autosomes.
    Using these settings, we generated three replicate datasets without selection
    and three replicate datasets with selection.
    Three additional datasets, used in Figure \ref{fig:homsap-dfe.constant},
    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $10,000$.
    Each of these simulations generated 100 diploid genomes.
    The configuration file used in combination with our \snakemake workflow
    for these simulations can be found at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_HomSap.yml}.

    \subsubsection*{Simulations of complete vaquita porpoise genomes}
    Simulation of vaquita porpoise genomes used 
    the two-epoch demographic model from \citet[\stdpopsim label \texttt{Vaquita2Epoch\_1R22}]{robinson2022critically}.
    Since no genetic map is currently available for this species, we assumed a constant recombination rate of  $10^{-8}$
    across all chromosomes \citep{morin2021}.
    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
    \citet[\stdpopsim label \texttt{Gamma\_R22}]{robinson2022critically}  to exons annotated from the vaquita genome assembly
    \citet[\stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}]{morin2021}.
    This DFE model implements a relationship between selection and dominance as follows:
    very deleterious mutations (with $s<-0.1$) are recessive ($h=0$),
    mutations with $s\in [-0.1,-0.01)$ have a dominance coefficient of $h=0.01$,
    mutations with $s\in [-0.01,-0.001)$ have a dominance coefficient of $h=0.1$,
    and very mild mutations (with $s\in [-0.001,0)$) are nearly additive ($h=0.4$).
    This discretized relationship matches that used by \textcite{robinson2022critically},
    and is easily implemented in \stdpopsim's \texttt{DFE} class.
    Each simulation generates 100 diploid genomes,
    with each genome consisting of 21 autosomes.
    Using these settings, we generated three replicate datasets without selection
    and three replicate datasets with selection.
    Three additional datasets, used in Figure \ref{fig:vaquita-dfe.constant},
    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $3,500$.
    \igcomment{If we end up using Fig. S3, then we should add a line describing it's simulation here.}
    The configuration file used in combination with our \snakemake workflow
    for these simulations can be found at 
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_PhoSin.yml}.

@@

    % ADK: I've added a section below that describes the DFE inference methods used in this study.
    For comparison among methods, estimates were rescaled. 
    As described above, each DFE-inference method scales selection coefficients slightly differently.
    We thus rescaled each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
    The output of \grapes  was scaled by a factor of  $\mu / \theta$,
    and the output of \polydfe was scaled by $2 \mu / \theta$, 
    where $\mu$ is the assumed mutation rate and 
    $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
    For \dadicli, the gamma distribution is parameterized using the shape and rate ($\alpha$ and $\nu$, respectively). 
    The mean value of the distribution, which is $s$ is transformed from $\alpha$ and $\nu$ and rescaled as $\alpha * \nu / (\theta / (4 \mu L p))$,
    where $L$ is the total number of sites available for the analysis,
    and $p$ is the proportion of sites under selection.
    Complete details for the DFE inference analysis can be found in the
    \snakemake workflow at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/dfe.snake}.

@@

    Each of the three sweep-detection methods is based on a test statistic.
    Reduced nucleotide diversity is based on nucleotide diversity  ($\pi$) 
    computed directly from the underlying tree sequences using \tskit \citep{ralph2020efficiently}
    over 10 equally sized and non-overlapping subwindows.
    \sweepfinder \citep{degiorgio2016sweepfinder2} is based on a composite likelihood ratio (CLR) 
    computed at 21 uniformly distributed positions along the simulated region.
    \diploshic \citep{kern2018diplos} uses supervised learning to train a sweep classifier.
    Its training phase used \texttt{discoal} \citep{kern2016discoal} to simulate 20 haploid sequences of length 1.21Mbp
    with a constant population size of $N_e=10,000$, a mutation rate of $2\times 10^{-8}$, and a recombination rate of $2\times 10^{-8}$.
    For \texttt{discoal} simulations that included selective sweeps,
    selection coefficients were drawn from a uniform distribution between $0.001$ and $0.10$,
    and the time of the sweep was chosen from a uniform distribution between $0$ and $0.01N_e$ generations in the past.
    For soft sweeps, the frequency at which selection was introduced was drawn from a uniform distribution between $0.0001$ and $0.1$.
    We note that since the training set of \diploshic was generated under a simple model with
    constant mutation and recombination rates and constant effective population size,
    we expect it to perform suboptimally on the simulated segments of human chromosome 1,
    which were generated using more realistic simulations (see \textbf{Simulations with selective sweeps} above).

    The three sweep-detection methods were associated with thresholds for their respective test statistics
    ($\pi$, \sweepfinder's CLR, or \diploshic's probability),
    which were determined using an empirical distribution representing a null model without sweeps.
    We considered two null models: a neutral evolution model (neutral),
    and a model with background selection from deleterious mutations in exons (BGS).
    Each null model was represented by a different set of $100,000$ simulations generated according to the appropriate model
    (see \textbf{Simulations with selective sweeps} above).
    Thus, for each test statistic we determined two thresholds corresponding to these two different null models (neutral and BGS)
    using the extreme $5\%$ quantile in each empirical distribution.
    Simulated regions where the central 1Mb had a recombination rate of less than $10^{-3}$cM were excluded from the analysis.
    This excluded 7 regions from the analysis. 
    Then, power (or true positive rate) to detect a sweep by a given method was determined for
    100 evenly-spaced positions on chromosome 1 as follows:
    each sweep-detection method was applied to the 1,000 
    simulated datasets generated with a hard sweep (neutral and BGS) for that position,
    and the power was set to be the proportion of simulations (out of 1,000)
    in which the test statistic exceeded the threshold determined by each of the two null models (neutral and BGS).
    We note that using an empirical distribution to determine the threshold for calling a selective
    sweep with \diploshic deviates from the original implementation of the method,
    however for consistency among methods we chose to take this approach.
    The complete \snakemake workflow for sweep detection can be found at
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/sweep_simulate.snake}.

@@

\section*{Funding}
    \label{funding}
    % we got too many people to do initials
    M.~Rodgrigues, S.~Tittes, N.~Pope, and A.~Kern were supported in part by NIH award R35GM148253.
    R.~Gutenkunst and L.~Tran were supported in part by NIH award R35GM149235.
    D.~Schrider was supported in part by NIH award R35GM138286.
    K.~Lohmueller was supported by NIH award R35GM119856.
    P.~Ralph, B.~Haller, A.~Kern, and B.~Jefferey
    were supported in part by NIH award R01HG012473.
    J.~Kelleher was supported by EPSRC grant EP/X024881/1
    and the Robertson Foundation.

\printbibliography

@@

%sweep power vs exon density
\begin{figure}
    \centering
    \includegraphics[width=0.8 \textwidth]{figures/sweeps/relationship_power_exon.pdf}
    \caption{
    %IG: copied new caption from fig 6 and adjusted.
        Power to detect selective sweeps as a function of local exon density.
        This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
        but with the 100 genomic segments sorted according to their
        exon density (percentage of exonic bases per window)instead of position along chromosome 1.
	    Genomic segments were simulated with sweeps under a three population out-of-Africa model
        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
        Three methods for detecting sweeps were applied to simulated data:
        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
        2) \diploshic \citep{kern2018diplos} (middle row),
        and 3) reduced diversity ($\pi$) (bottom row).
        Power (true positive rate) is shown for these methods for the CEU and YRI
        samples (left and right respectively).
        The thresholds of the test statistics were set to obtain a
        $5\%$ false positive rate under a neutral null model (blue)
        and a null model with background selection from deleterious mutations in exons (red).
        Fitted lines represent loess smoothed regressions.
    }
    \label{fig:power-exon}
\end{figure}
